---
title: "Aides Drug"
author: "Areli Nava"
date: "6/21/23"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
I need to use a filter demand form the helper cheat sheet. The filter to get the data out the triUmp that is comprise of drug users. I will proceed to use a set of codes later on in this process.



```{r}
DRUGUSR<-filter(triUmp,DrugUser=="yes")
```

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

```{r}
DRUGUSR <- DRUGUSR %>%
mutate(across(where(is.character), as.factor))
```


## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->






#### Plot barplots ggplot way

```{r}
dd2 <- DRUGUSR %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



The first plot shows how many people got infected. It looks like there may be less than 100 people with AIDS.  In the % plot it treats both groups in % terms. 25% of people that had AIDS has taken the drug. 70% of the of the people that has AIDS go the placebo. more people on the placebo side got sick because they didn't take the drug. it prevents those who took it from getting AIDs.



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=DRUGUSR)
rowPerc(table1)
colPerc(table1)
```

Out of the people who got AIDS 24.29% have taken the drug. &5.71% have taken the placebo. The people that havent taken anything is 47.59% to 52.41%. this signals that a drug is working.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the probability becomes slightly low to the number of 0000000423. this tells us that we reject the hypothesis and explains the drug works. the small p value explains something is occurring.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

With this project we learned to incorporate the Mytricks.rmd and the dplyr. We observed in a graph how the drug and the placebo take into affect. The drugs were taking positive effects and worked on the people. The inferential test concluded that the drug prevented AIDS and worked.
